### 2017-01-23 thesis/intro/intro.tex
Finished editing introduction - tmp commit

\noindent
Many of the hallmarks are acquired by the cancer cells during tumout growth, through many genomic changes and mutations of the essential genes that allow cancer cells to survive, proliferate, invade and metastasise.
Thus, it is only logical to induce that many, if not all, cancers are more prone to genetic mutations and restructuring, which allows the cancers to subsequently acquire the already described Hallmarks of Cancer \citep{Hanahan2011}.

@@

Oncogenes are genes that promote or ``accelerate'' the process of tumorigenesis and/or tumour progression, whereas tumour suppressor genes prevent or ``decelerate'' those effects \citep{Vogelstein2004a}.
Cancers usually mutate oncogenes in such a way that they become constitutively activated, and mutate the tumour suppressor genes so that these genes are inactivated and/or have reduced expression \citep{Vogelstein2004a}.
Another thing to note for these genes is that oncogenes usually only require a ``one-hit'' mutation that affects a single allele, whereas tumour suppressor genes require mutations in both maternal and paternal alleles to have any effect \citep{Stratton2009,Vogelstein2004a}.

Stability genes, also known as caretakers, are genes that are involved in, for example,  \gls{mmr}, \gls{ner} and \gls{ber}; in other words, genes that are involved in \acrshort{dna} repair mechanisms \citep{Vogelstein2004a}.
These caretaker genes in effect prevents the cell to develop new mutations from \acrshort{dna} replication and mitosis.
However, only the oncogenes or the tumour suppressor genes are able to directly affect the biology of the cell, and therefore mutations in the stability genes affect the mutation rates of the other classes of genes, ultimately leading to tumorigenesis \citep{Vogelstein2004a}.
\\

\noindent
There are two types of mutations that occurs in an individual: germline mutation and somatic mutation.
Germline mutations are the mutations that are passed on from your parents, and somatic mutations are the mutations acquired by the cell over the course of the individual's lifetime.

Unfortunately, there are germline mutations that make the individual more prediposed to cancer \citep{Vogelstein2004a}.
These individuals have a mutation in one or more of the classes of cancer-related genes, which makes them more likely to develop tumour than their normal pairs \citep{Vogelstein2004a}.
As a consequence, these individuals may develop multiple tumours at an early stage of their lives than those who develop tumours from somatic mutations alone, as these individuals already have a mutation that allows all of the cells in their body to have one or more of the Hallmarks of Cancers \citep{Vogelstein2004a}.

In contrast, somatic mutations are acquired during an individual's lifetime and is not passed down to their children.
Somatic mutations can either be a driver or a passenger mutation, where a driver mutation helps tumorigenesis and its subsequent progression (in other words, mutations in oncogene, tumour suppressor, or stability gene), and passenger mutation is where the mutation has no selective advantage for the cancer and does not contribute to tumour progression \citep{Stratton2009}.
Since cancers require multiple driver mutations, somatic mutations are an essential part of tumour development.
Perhaps that is one of the reasons why, in some cancers, germline mutation in the stability genes cause greater susceptibility to cancers than mutations in oncogene or tumour suppressor gene \citep{Vogelstein2004a}.

@@

For obesity, since there is a clear evidence of the cause (excessive eating and lack of exercise) and a well defined mechanism of how the disease is developed (excess calories are converted into fat), it is relatively easy to speculate the effect of the environment on the disease (for example, abundance of high calory food).
On the other hand, cancers are developed through many cellular pathways (Hallmarks of Cancers) which are not fully uncovered yet, with either too many or very few causes associated with those pathways (for example, mutations, inflammation and/or growth factors), and therefore it is difficult to assess how, rather than whether, a given environmental factor causes specific cancer.

With that said, there has been a few environmental factors that have a strong evidence of causing cancer, though the complete mechanism of action for some of these must be elucidated in the coming years.
Out of these, smoking, \gls{uv} radiation, and diet and obesity will be covered as examples.
\\

% Smoking
\noindent
Smoking is a known risk factor for many cancers; the obvious one being lung cancer \citep{Gandini2008,Hecht1999}.
For current smokers, the relative risk is greater than 1.5 in many cancers, including liver, cervix and kidney, and it can rise up to 8.96 in lung cancer \citep{Gandini2008}.
Even if smokers stop smoking, the relative risk does not decrease until the first 5 years have passed and will never return to the level of non-smokers \citep{Hecht1999}.

In tobacco smoke, there are literally hundreds and thousands of chemical compounds, of which about 20 compounds have been shown to have carcinogenic effects \citep{Hecht1999}.
Among these, \glspl{pah} and \gls{nnk} have been shown to cause lung cancer in mice when administered systemically \citep{Hecht1999}.
These carcinogens are processed first by \gls{cyp} enzymes (termed meta\-bolic activation), and the product can either be processed further to be detoxified completely and cause no harm to the cell, or alternatively, the product can covalently bind to the \acrshort{dna}, forming \acrshort{dna} adducts \citep{Hecht1999}.
Unless these adducts are resolved correctly by the \acrshort{dna} repair mechanisms, it can result in a permanent mutation within the genome and lead to oncogenesis \citep{Hecht1999}.

In addition to this, cigarettes contain free radicals and chemicals that cause tremendous oxidative stress and damage to the cells \citep{Hecht1999}.
These radicals cause \acrshort{dna} nicks and damages the \acrshort{dna}, again requring the cell's \acrshort{dna} repair mechanisms to remove the damage so it does not become permanent.
\\

% UV radiation
\noindent
Another important environmental factor to consider is \gls{uv} radiation.
Exposure to \gls{uv} radiation has been significantly associated with the increased risks of skin cancers \citep{Armstrong2001,Gallagher2006}.
The risks of developing any skin cancers can vary between race (Caucasian people are more likely to get skin cancers), and potentially the age when they were exposed to high amounts of \gls{uv} radiation \citep{Armstrong2001,Gallagher2006}.

Again, the mechanism of how \gls{uv} radiation may cause  skin cancers is related to the \acrshort{dna} repair system.
\gls{uv} radiation causes the formation of a dimer in the adjacent pyrimidines, resulting in a covalent bond formation between these nucleotides \citep{Friedberg2003,Hoeijmakers2001}.
These dimers introduce bulky structures in the \acrshort{dna} that prevent the \acrshort{dna} replication and transcription enzymes from functioning properly, and if they are not repaired, it may cause lasting mutations \citep{Friedberg2003,Hoeijmakers2001}.
These damages caused by \gls{uv} radiation can be relieved by the \gls{ner} system or the photoreactivation process, where a photoreactivating enzyme uses light to monomerize the dimeric pyrimidines \citep{Friedberg2003}.
\\

% Diet and obesity
\noindent
Finally, diet and obesity are known risk factors for many types of cancers \citep{Ames1995a,Calle2004}.
Though more will be covered in details in \cref{sec:obesity_and_cancer}, the effect of obesity on cancers will be briefly mentioned here.

Epidemiological evidence of obesity being a risk factor for many cancers has been well established, yet the mechanisms of its contribution to cancer is still  speculative \citep{Calle2003,Kelesidis2006a}.
Although there are a lot of possible mechcanisms, there is growing evidence of the role of chronic inflammation caused by increased fat mass in the body \citep{Kelesidis2006a,Lumeng2011,Hernandez2013}.
Free fatty acids from excess adipose tissue and/or diet can be recognised by the \gls{tlr}4, which can then activate \gls{nfkb} \citep{Lumeng2011}.
Clearly, the result of the activation of \gls{nfkb} can lead to the activation of one of the Hallmarks of Cancers -- tumour-promoting inflammation.

The nutritional content of the food that people consume is also important, as it can both directly and indirectly influence the likelihood of tumorigenesis, as well as its progression \citep{Ames1995a}.
Many fruits and vegetables contain antioxidants as well as essential viatmins and minerals.
Antioxidants are able to neutralise the effect of \gls{ros},  and many viatmins have been associated with reduction of many types of cancers \citep{Ames1995a}.
In fact, ascorbic acid (vitamin C) is able to prevent the lipid peroxidation of lipid plasma caused by tobacco smokes \textit{in vitro}, which could be why smokers, compared to non-smokers, have decreased levels of ascorbic acid \citep{Hecht1999}.

By maintaining healthy diet, it may be possible to reduce the risks of many cancers and obesity, as well as the comorbidities associated with them.
\\

\noindent
As one can imagine, \acrshort{dna} damages caused by smoking and/or \gls{uv} radiation have the potential to damage the genome and initiate cancer if wrong genes are mutated, such as p53 tumour suppressor gene.
Healthy diet is also crucial for lowering the risk of cancers through beneficial vitamins and antioxidants, and prevent obesity as well as chronic inflammation that accompanies with it.
There are many other environmental factors associated with cancers, and limiting the exposure to detrimental environmental factors is crucial for the prevention of any types of cancers, as well as the more specific types of cancers associated with these factors, such as lung and skin cancers.
\\

@@

The RAS-RAF-MEK-ERK signalling pathway is a pathway involved in cell survival, growth and proliferation \citep{Samatar2014,Wilhelm2006}.
The members of this pathway are often mutated in many types of cancers (approximately a third of cancers contain a mutation in the RAS oncogene), causing an overactivation of this pathway \citep{Samatar2014}.
Development of drugs that target  the downstream RAF and its mutant forms (such as BRAF$^{V600E}$) have been successful, and many RAF inhibitor drugs such as sorafenib and vemurafenib have been approved for clinical use \citep{Samatar2014,Wilhelm2006}.
Even though the signalling pathways are well understood and the quality of the drugs that are developed based on these knowledge is improving, there is always the possibility of tumour to acquire resistance to those drugs \citep{Samatar2014}.

In fact, there has been evidence of pathway cross-talks between RAS-ERK pathway and \gls{pi3k}/Akt pathway, which may help the tumour cells achieve resistance against those drugs \citep{Moelling2002,Zimmermann1999a}.
A study in a \textit{Drosophila} model showed that the inhibition of just a single pathway does not necessarily stop the effect of the pathway, as the signal can flow through alternate pathways \citep{Dar2012}.
Pathway cross-talk and increased flux through alternate pathways could be one of the ways in which tumour cells acquire resistance to drugs.
However, the effectiveness of targeting multiple pathways in humans is yet to be confirmed and further investigations are required to solidify the mode of action and its relevance to tumour biology.
\\

\noindent
There are many difficulties accompanied by cancer treatment such as resistance to drugs and metastasis, and there is always a need for new ways to treat cancer patients.
Further understanding of the tumour biology, its mechanism of resistance to drugs, and the effect of targeting multiple pathways (or targeting effective pathways for a patient) will definitely become important in the near future, both for novel drug development and better clinical treatment plan for individual patients.

@@

As mentioned briefly in \cref{sub:causes_and_risk_factors_of_cancer}, obesity is considered a major risk factor for many types of cancers.
The fact that the prevalence of obesity is extremely high, it is important to establish the link between the apparent risks associated with cancer.

@@

Although the mechanisms in which obesity causes cancer is still under debate, the association between obesity and cancer has long been established.

Perhaps the most convincing evidence of the association between obesity and cancer was first presented by \citet{Calle2003}.
In this study, the link between \gls{bmi} and variety of cancers were investigated in 900,000 adults between 1982 and 1998 \citep{Calle2003}.
The results from this study showed that morbidly obese men and women (\gls{bmi} \textgreater{} 40) had 52\% and 62\% higher chance of dying from all cancers compared to normal weight adults (18.5 \textgreater{} \gls{bmi} \textgreater{} 25), respectively \citep{Calle2003}.
\gls{bmi} was also associated with higher rates of death from esophageal, colorectal, liver, gallbladder, pancreas, kidney, and some sex-specific cancers \citep{Calle2003}.
Furthermore, the death rates had a positive linear correlation with increasing \gls{bmi} in all cancers \citep{Calle2003}.

In a more recent meta-analysis study by \citet{Renehan2008} showed that an increase \gls{bmi} by 5 kg/m$^2$ significantly increased the risks of esophageal, thyroid, colon and renal cancers in men; endometrial, gallbladder, esophageal and renal cancers in women.
The summary of the risks associated with various cancers from this study are shown in \cref{tab:renehan_cancer_risks}.
This study showed that the association between \gls{bmi} and cancers were sex-specific, applicable to wide range of cancer types, and generally consistent across geographic populations \citep{Renehan2008,Roberts2010}.

Taken together, it is undeniable that obesity is a major risk factor for many types of cancers, and it may even be the cause for some of these cancers.

\begin{table}[h]
	\centering
	\begin{threeparttable}
		\caption{Summary of the risk ratios\tnote{1} associated with each type of cancers from the meta-analysis study by \citet{Renehan2008}\tnote{2}.}
		\label{tab:renehan_cancer_risks}
		\begin{tabular}{lcc}
													   & Men               & Women\\
			\hline
			\rule{0pt}{2.25ex}Colon                 & 1.24 (1.20, 1.28) & 1.09 (1.05, 1.13)\\
			Rectum                                  & 1.09 (1.06, 1.12) & NA\tnote{3}\\
			Gallbladder                             & NA                & 1.59 (1.02, 2.47)\\
			Leukemia                                & 1.08 (1.02, 1.14) & 1.17 (1.04, 1.32)\\
			Malignant melanoma                      & 1.17 (1.05, 1.30) & NA\\
			Multiple myeloma                        & 1.11 (1.05, 1.18) & 1.11 (1.07, 1.15)\\
			Non-Hodgkin lymphoma                    & 1.06 (1.03, 1.09) & 1.07 (1.00, 1.14)\\
			Esophageal adenocarcinoma               & 1.52 (1.33, 1.74) & 1.51 (1.31, 1.74)\\
			Pancreatic                              & NA                & 1.12 (1.02, 1.22)\\
			Renal                                   & 1.24 (1.15, 1.34) & 1.34 (1.25, 1.43)\\
			Thyroid                                 & 1.33 (1.04, 1.70) & 1.14 (1.06, 1.23)\\
			Prostate                                & 1.03 (1.00, 1.09) & NA\\
			Post-menopausal breast                  & NA                & 1.12 (1.08, 1.16)\\
			Endometrial (\textless{}27 kg/m$^2$)    & NA                & 1.221 (1.084, 1.376)\\
			Endometrial (\textgreater{}27 kg/m$^2$) & NA                & 1.729 (1.598, 1.872)\\
			\hline
			\hline
		\end{tabular}
		\begin{tablenotes}
			\begin{footnotesize}
				\item [1] All risk ratios are per increase in 5 kg/m$^2$ \gls{bmi}; 95\% confidence intervals in brackets.
				\item [2] Table adapted from \citet{Roberts2010} paper, where the original data in the table was taken from \citet{Renehan2008} study.
				\item [3] Not applicable.
			\end{footnotesize}
		\end{tablenotes}
	\end{threeparttable}
\end{table}

@@

The exact mechanisms in which obesity causes cancer have still not been fully elucidated yet.
There are several hypotheses that links obesity with cancers: disruption of various hormone levels, chronic and sustained inflammation, and activation of the insulin/\gls{igf} pathway.

Previously thought to be a group of cells where the body could store and manage excess fats, it has recently become apparent that adipose tissue is also a metabolically active organ that has a significant role in the endocrine system \citep{Roberts2010}.
Increased adiposity is known to influence the bioavailability of endogenous sex hormones, including estrogens, androgens and progesterone \citep{Calle2004}.
\Glspl{adipocyte} synthesise enzymes, such as aromatase, that convert precursor molecules into different forms of androgens and estrogens \citep{Calle2004}.
Many studies have confirmed the mitogenic effect of sex hormones on certain cell types, mainly breast and endometrial cells \citep{Roberts2010}.

On top of the effects that \glspl{adipocyte} have on circulating concentration of sex hormones, they also produce adipose tissue-specific hormones called adipokines \citep{Roberts2010}.
Of the many adipokines identified so far, leptin and adipo\-nectin have been studied the most in terms of cancer development \citep{Renehan2006,Roberts2010}.
Leptin is known to activate not only the satiety signalling pathway, but also the \gls{jak}/\gls{stat}, \gls{pi3k}/Akt signalling pathways \citep{Garofalo2006,Renehan2006}.
Activation of these pathways lead to the activation of their target genes that are involved in cell proliferation and cell survival, as well as pro-angiogenic factors \citep{Garofalo2006}.

Adiponectin is a hormone that is synthesised and secreted only by the \glspl{adipocyte} \citep{Kelesidis2006a}.
Unlike leptin, where the concentration of leptin is higher in obese people, adiponectin levels are lower in obese people compared to the normal weight people \citep{Kelesidis2006a,Renehan2006}.
Adiponectin has anti-angiogenic and anti-proliferative effect on the cells, and is secreted by mature adipocytes but not by the premature pre-adipocytes \citep{Gilbert2013}.
During chronic inflammatory state in obese patients, pro-inflammatory proteins such as \gls{tgfb} and \gls{ifny} can block the maturation of the pre-adipocytes, thus decreasing the level of adiponectin and the accompanying benefecial effects \citep{Gilbert2013}.

Alternatively, or possibly in conjunction with hormone disruption, inflammation caused by obesity may be associated with tumour-promoting environment.
\Gls{tlr}4 is a type of \gls{prr} and, as mentioned earlier, it is able to recognise circulating free fatty acids and trigger immune response via \gls{nfkb} pathway \citep{Lumeng2011}.
In addition to the \gls{nfkb} pathway, recent evidence have shown that \gls{tlr}4 could also activate the \gls{jak}/\gls{stat3} pathway \citep{Yu2014}.
Activation of these pathways will further enhance the development of the tumour-promoting microenvironment caused by these unwanted inflammatory responses of obesity.
Furthermore, since it is likely that obese people have less adiponectin concentration, obesity-induced inflammation may affect these patients more than those with normal weight.

Finally, insulin/\gls{igf} pathway may have a major role in encouraging tumorigenesis in obese patients.
\Gls{igf} system involves two ligands (\gls{igf}-I and \gls{igf}-II), its respective receptors (\gls{igf}-IR and \gls{igf}-IIR), and six binding proteins (\gls{igfbp} 1 to 6) \citep{Roberts2010}.
\gls{igf} signalling results in the activation of \gls{pi3k}/Akt and \gls{erk} pathways, resulting in cell proliferation and reduced apoptosis \citep{Roberts2010}.
Though insulin is also capable of activating these pathways, it is thought that the mitogenic effects are controlled mainly through the \gls{igf} receptors \citep{Roberts2010}.

>>>>

\noindent
Many of the hallmarks are acquired by the cancer cells during tumour growth, through many genomic changes and mutations of the essential genes that allow cancer cells to survive, proliferate, invade and metastasise.
Thus, it is only logical to induce that many, if not all, cancers are more prone to genetic mutations and restructuring, which allows the cancers to subsequently acquire the already described Hallmarks of Cancer \citep{Hanahan2011}.

@@

Oncogenes are genes that promote or ``accelerate'' the process of tumorigenesis and/or tumour progression, whereas tumour suppressor genes prevent or ``decelerate'' those effects \citep{Vogelstein2004a}.
Cancers usually mutate oncogenes in such a way that they become constitutively activated and mutate the tumour suppressor genes so that these genes are inactivated and/or have reduced expression \citep{Vogelstein2004a}.
Another thing to note for these genes is that oncogenes usually only require a ``one-hit'' mutation that affects a single allele, whereas tumour suppressor genes generally require mutations in both maternal and paternal alleles to have any effect \citep{Stratton2009,Vogelstein2004a}.

Stability genes, also known as caretakers, are genes that are involved in, for example,  \gls{mmr}, \gls{ner} and \gls{ber}; in other words, genes that are involved in \acrshort{dna} repair mechanisms \citep{Vogelstein2004a}.
These caretaker genes help prevent the cell from developing new mutations during \acrshort{dna} replication and mitosis.
However, only the oncogenes or the tumour suppressor genes are able to directly affect the biology of the cell, and therefore mutations in stability genes tend to affect the mutation rates of the other classes of genes, ultimately leading to tumorigenesis \citep{Vogelstein2004a}.

There are two types of mutations that occurs in an individual: germline mutation and somatic mutation.
Germline mutations are the mutations that are passed on from an individual's parents, and somatic mutations are the mutations acquired by the cell over the course of the individual's lifetime.

Unfortunately, there are germline mutations that make the individual more prediposed to cancer \citep{Vogelstein2004a}.
These individuals have a mutation in one or more of the classes of cancer-related genes, which makes them more likely to develop a tumour than individuals without the mutation \citep{Vogelstein2004a}.
In contrast, somatic mutations are acquired during an individual's lifetime and is not passed down to their children.
Somatic mutations can either be a driver or a passenger mutation, where a driver mutation helps tumorigenesis and its subsequent progression (in other words, mutations in oncogene, tumour suppressor, or stability gene), and passenger mutation have neither a selective advantage nor contribute to tumour progression \citep{Stratton2009}.

Since cancers require multiple driver mutations, somatic mutations are an essential part of tumour development.
Perhaps that is one of the reasons why, in some cancers, germline mutation in the stability genes cause greater susceptibility to cancers than mutations in oncogene or tumour suppressor gene \citep{Vogelstein2004a}.

@@

% For obesity, since there is a clear evidence of the cause (excessive eating and lack of exercise) and a well defined mechanism of how the disease is developed (excess calories are converted into fat), it is relatively easy to speculate the effect of the environment on the disease (for example, abundance of high calory food).
Cancers are developed through many cellular pathways (Hallmarks of Cancers) which are not fully uncovered yet, with either too many or very few causes associated with those pathways.
Therefore it is difficult to assess how, rather than whether, a given environmental factor causes specific cancer.
With that said, several environmental factors have a strong evidence of causing cancer.
% , though the complete mechanism of action for some of these must be elucidated in the coming years.
% Out of these diet and obesity will be covered as an example.

Smoking is a known risk factor for many cancers; the obvious one being lung cancer \citep{Gandini2008,Hecht1999}.
For current smokers, the relative risk is greater than 1.5 in many cancers, including liver, cervix and kidney, and it can rise up to 8.96 in lung cancer \citep{Gandini2008}.
% Even if smokers stop smoking, the relative risk does not decrease until the first 5 years have passed and will never return to the level of non-smokers \citep{Hecht1999}.
% In tobacco smoke, there are literally hundreds and thousands of chemical compounds, of which about 20 compounds have been shown to have carcinogenic effects \citep{Hecht1999}.
% Among these, \glspl{pah} and \gls{nnk} have been shown to cause lung cancer in mice when administered systemically \citep{Hecht1999}.
Tobacco smoke contain many carcinogens which are processed first by \gls{cyp} enzymes (termed meta\-bolic activation), and the product can either be processed further to be detoxified completely and cause no harm to the cell, or alternatively, the product can covalently bind to the \acrshort{dna}, forming \acrshort{dna} adducts \citep{Hecht1999}.
Unless these adducts are resolved correctly by the \acrshort{dna} repair mechanisms, it can result in a permanent mutation within the genome and lead to oncogenesis \citep{Hecht1999}.
% In addition to this, cigarettes contain free radicals and chemicals that cause tremendous oxidative stress and damage to the cells \citep{Hecht1999}.
% These radicals cause \acrshort{dna} nicks and damages the \acrshort{dna}, again requring the cell's \acrshort{dna} repair mechanisms to remove the damage so it does not become permanent.

% Another important environmental factor to consider is \gls{uv} radiation.
Exposure to \gls{uv} radiation has been significantly associated with the increased risks of skin cancers \citep{Armstrong2001,Gallagher2006}.
% The risks of developing any skin cancers can vary between race (Caucasian people are more likely to get skin cancers), and potentially the age when they were exposed to high amounts of \gls{uv} radiation \citep{Armstrong2001,Gallagher2006}.
Similar to smoking, the mechanism of how \gls{uv} radiation may cause  skin cancers is related to the \acrshort{dna} repair system.
\gls{uv} radiation causes the formation of a dimer in the adjacent pyrimidines, resulting in a covalent bond formation between these nucleotides \citep{Friedberg2003,Hoeijmakers2001}.
These dimers introduce bulky structures in the \acrshort{dna} that prevent the \acrshort{dna} replication and transcription enzymes from functioning properly, and it may cause lasting mutations if not repaired \citep{Friedberg2003,Hoeijmakers2001}.
These damages caused by \gls{uv} radiation can be relieved by the \gls{ner} system or the photoreactivation process, where a photoreactivating enzyme uses light to monomerize the dimeric pyrimidines \citep{Friedberg2003}.

Diet and obesity are known risk factors for many types of cancers as well \citep{Ames1995a,Calle2004}.
% Though more will be covered in details in \cref{sec:obesity_and_cancer}, the effect of obesity on cancers will be briefly mentioned here.
Epidemiological evidence of obesity being a risk factor for many cancers has been well established, yet the mechanisms of its contribution to cancer is still  speculative \citep{Calle2003,Kelesidis2006a}.
Although there are a lot of possible mechcanisms, there is growing evidence of the role of chronic inflammation caused by increased fat mass in the body \citep{Kelesidis2006a,Lumeng2011,Hernandez2013}.
Free fatty acids from excess adipose tissue and/or diet can be recognised by \gls{tlr} (specifically \gls{tlr}4), which can then activate \gls{nfkb}, leading to the activation of one of the Hallmarks of Cancer -- tumour-promoting inflammation \citep{Lumeng2011}.
% The nutritional content of the food that people consume is also important, as it can both directly and indirectly influence the likelihood of tumorigenesis, as well as its progression \citep{Ames1995a}.
% Many fruits and vegetables contain antioxidants as well as essential viatmins and minerals.
% Antioxidants are able to neutralise the effect of \gls{ros},  and many viatmins have been associated with reduction of many types of cancers \citep{Ames1995a}.
% In fact, ascorbic acid (vitamin C) is able to prevent the lipid peroxidation of lipid plasma caused by tobacco smokes \textit{in vitro}, which could be why smokers, compared to non-smokers, have decreased levels of ascorbic acid \citep{Hecht1999}.
% By maintaining healthy diet, it may be possible to reduce the risks of many cancers and obesity, as well as the comorbidities associated with them.

% As one can imagine, \acrshort{dna} damages caused by smoking and/or \gls{uv} radiation have the potential to damage the genome and initiate cancer if wrong genes are mutated, such as p53 tumour suppressor gene.
% Healthy diet is also crucial for lowering the risk of cancers through beneficial vitamins and antioxidants, and prevent obesity as well as chronic inflammation that accompanies with it.
There are many other environmental factors associated with cancers, and limiting the exposure to detrimental environmental factors is crucial for the prevention of any type of cancers, as well as the more specific type of cancers associated with these factors, such as lung and skin cancers.

@@

The RAS-RAF-MEK-ERK signalling pathway is  involved in cell survival, growth and proliferation \citep{Samatar2014,Wilhelm2006}.
The members of this pathway are often mutated in many types of cancer (approximately a third of cancers contain a mutation in the RAS oncogene), causing an overactivation of this pathway \citep{Samatar2014}.
Development of drugs that target  the downstream RAF and its mutant forms (such as BRAF$^{V600E}$) have been successful, and many RAF inhibitor drugs such as sorafenib and vemurafenib have been approved for clinical use \citep{Samatar2014,Wilhelm2006}.
Even though the signalling pathways are well understood and the quality of the drugs that are developed based on these knowledge is improving, there is always the possibility of tumour to acquire resistance to those drugs \citep{Samatar2014}.

In fact, there has been evidence of pathway cross-talks between RAS-ERK pathway and \gls{pi3k}/Akt pathway, which may help the tumour cells achieve resistance against those drugs \citep{Moelling2002,Zimmermann1999a}.
A study in a \textit{Drosophila} model showed that the inhibition of just a single pathway does not necessarily stop the effect of the pathway, as the signal can flow through alternate pathways \citep{Dar2012}.
Pathway cross-talk and increased flux through alternate pathways could be one of the ways in which tumour cells acquire resistance to drugs.
However, the effectiveness of targeting multiple pathways in humans is yet to be confirmed and further investigations are required to solidify the mode of action and its relevance to tumour biology.

Further understanding of the tumour biology, its mechanism of resistance to drugs, and the effect of targeting multiple pathways (or targeting effective pathways for a patient) will definitely become important in the near future, both for novel drug development and better clinical treatment plan for individual patients.

@@

As mentioned briefly in \cref{sub:causes_and_risk_factors_of_cancer}, obesity is considered a major risk factor for many types of cancer.
The fact that the prevalence of obesity is extremely high, it is important to establish the link between the apparent risks associated with cancer.

@@

Although the mechanisms in which obesity causes cancer is still under debate, the association between obesity and cancer has long been established.
Perhaps the most convincing evidence of the association between obesity and cancer was first presented by \citet{Calle2003}.
In this study, the link between \gls{bmi} and variety of cancers were investigated in 900,000 adults between 1982 and 1998 \citep{Calle2003}.
The results from this study showed that morbidly obese men and women (\gls{bmi} \textgreater{} 40) had 52\% and 62\% higher chance of dying from all cancers compared to normal weight adults (18.5 \textgreater{} \gls{bmi} \textgreater{} 25), respectively \citep{Calle2003}.
\gls{bmi} was also associated with higher rates of death from esophageal, colorectal, liver, gallbladder, pancreas, kidney, and some sex-specific cancers \citep{Calle2003}.
Furthermore, the death rates had a positive linear correlation with increasing \gls{bmi} in all cancers \citep{Calle2003}.

In a more recent meta-analysis study, \citet{Renehan2008} showed that an increase in \gls{bmi} by 5 kg/m$^2$ significantly increased the risks of esophageal, thyroid, colon and renal cancers in men, and endometrial, gallbladder, esophageal and renal cancers in women.
The summary of the risks associated with various cancers from this study are shown in \cref{tab:renehan_cancer_risks}.
This meant that the association between \gls{bmi} and cancers were sex-specific, applicable to wide range of cancer types, and generally consistent across different geographic populations \citep{Renehan2008,Roberts2010}.
It is undeniable that obesity is a major risk factor for many types of cancer and it may even be the cause for some of these cancers.

% Taken together, it is undeniable that obesity is a major risk factor for many types of cancers, and it may even be the cause for some of these cancers.

\begin{table}[htb]
	\centering
	\begin{threeparttable}
		\caption{Summary of the risk ratios\tnote{1} associated with each type of cancers from the meta-analysis study by \citet{Renehan2008}\tnote{2}.}
		\label{tab:renehan_cancer_risks}
		\begin{tabular}{lcc}
													   & Men               & Women\\
			\hline
			\rule{0pt}{2.25ex}Colon                 & 1.24 (1.20, 1.28) & 1.09 (1.05, 1.13)\\
			Rectum                                  & 1.09 (1.06, 1.12) & NA\tnote{3}\\
			Gallbladder                             & NA                & 1.59 (1.02, 2.47)\\
			Leukemia                                & 1.08 (1.02, 1.14) & 1.17 (1.04, 1.32)\\
			Malignant melanoma                      & 1.17 (1.05, 1.30) & NA\\
			Multiple myeloma                        & 1.11 (1.05, 1.18) & 1.11 (1.07, 1.15)\\
			Non-Hodgkin lymphoma                    & 1.06 (1.03, 1.09) & 1.07 (1.00, 1.14)\\
			Esophageal adenocarcinoma               & 1.52 (1.33, 1.74) & 1.51 (1.31, 1.74)\\
			Pancreatic                              & NA                & 1.12 (1.02, 1.22)\\
			Renal                                   & 1.24 (1.15, 1.34) & 1.34 (1.25, 1.43)\\
			Thyroid                                 & 1.33 (1.04, 1.70) & 1.14 (1.06, 1.23)\\
			Prostate                                & 1.03 (1.00, 1.09) & NA\\
			Post-menopausal breast                  & NA                & 1.12 (1.08, 1.16)\\
			Endometrial (\textless{}27 kg/m$^2$)    & NA                & 1.221 (1.084, 1.376)\\
			Endometrial (\textgreater{}27 kg/m$^2$) & NA                & 1.729 (1.598, 1.872)\\
			\hline
			\hline
		\end{tabular}
		\begin{tablenotes}
			\begin{footnotesize}
				\item [1] All risk ratios are per increase in 5 kg/m$^2$ \gls{bmi}; 95\% confidence intervals in brackets.
				\item [2] Table adapted from \citet{Roberts2010} paper, where the original data in the table was taken from \citet{Renehan2008} study.
				\item [3] Not applicable.
			\end{footnotesize}
		\end{tablenotes}
	\end{threeparttable}
\end{table}

@@

The exact mechanisms in which obesity causes cancer have still not been fully elucidated yet.
There are several hypotheses that link obesity with cancer: disruption of various hormone levels, chronic and sustained inflammation, and activation of the insulin/\gls{igf} pathway.

% Previously thought to be a group of cells where the body could store and manage excess fats, it has recently become apparent that adipose tissue is also a metabolically active organ that has a significant role in the endocrine system \citep{Roberts2010}.
Adipose tissue has long been thought to be a group of cells to store excess fats, but it has recently become apparent that adipose tissue is also a metabolically active endocrine organ \citep{Roberts2010}.
Increased adiposity is known to influence the bioavailability of endogenous sex hormones, including estrogens, androgens and progesterone \citep{Calle2004}.
\Glspl{adipocyte} synthesise enzymes, such as aromatase, that convert precursor molecules into different forms of androgens and estrogens \citep{Calle2004}.
Many studies have confirmed the mitogenic effect of sex hormones on certain cell types, mainly breast and endometrial cells \citep{Roberts2010}.

On top of the effects that \glspl{adipocyte} have on circulating concentration of sex hormones, they also produce adipose tissue-specific hormones called adipokines \citep{Roberts2010}.
Of the many adipokines identified so far, leptin and adipo\-nectin have been studied the most in terms of cancer development \citep{Renehan2006,Roberts2010}.
Leptin is known to activate not only the satiety signalling pathway, but also the \gls{jak}/\gls{stat} and \gls{pi3k}/Akt signalling pathways \citep{Garofalo2006,Renehan2006}.
Activation of these pathways lead to the activation of target genes that are involved in cell proliferation and cell survival, as well as pro-angiogenic factors \citep{Garofalo2006}.

Adiponectin is a hormone that is synthesised and secreted only by the \glspl{adipocyte} \citep{Kelesidis2006a}.
Unlike leptin, where the concentration of leptin is higher in obese people, adiponectin levels are lower in obese people compared to the normal weight people \citep{Kelesidis2006a,Renehan2006}.
Adiponectin has anti-angiogenic and anti-proliferative effect on the cells, and is secreted by mature adipocytes but not by the premature pre-adipocytes \citep{Gilbert2013}.
During chronic inflammatory state in obese patients, pro-inflammatory proteins such as \gls{tgfb} and \gls{ifny} can block the maturation of the pre-adipocytes, thus decreasing the level of adiponectin and the accompanying beneficial effects \citep{Gilbert2013}.

Alternatively, or possibly in conjunction with hormone disruption, inflammation caused by obesity may be associated with tumour-promoting environment.
\Gls{tlr}4 is a type of \gls{prr} and, as mentioned earlier, is able to recognise circulating free fatty acids and trigger immune response via \gls{nfkb} pathway \citep{Lumeng2011}.
In addition to \gls{nfkb} pathway, recent evidence have shown that \gls{tlr}4 could also activate the \gls{jak}/\gls{stat3} pathway \citep{Yu2014}.
Activation of these pathways will further enhance the development of the tumour-promoting microenvironment caused by these unwanted inflammatory responses of obesity.
Furthermore, since it is likely that obese people have less adiponectin concentration, obesity-induced inflammation may affect these patients more than those with normal weight.
